Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-10-06 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:10:06	16:04:07	(Tue 6.Oct)

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
% 
% \begin{itemize}
% \item Dynamic weights $\Rightarrow$ greater contrast/resolution potential
% \item But:
% \begin{itemize}
% \item Active system $\Rightarrow$ noise and signal correlated $\Rightarrow$ adaptive beamformers no longer robust.
% \item Subarray averaging decorrelate the noise, but in the process we sacrifice resolution.
% \item Computationally complex O($M^3$) due to inversion of a spatial covariance matrix. LCA is only O($MW$). Even if Capon subspace/beamspace versions exist, they remain more complex than the LCA, and is not as ideally suited for implementation of SIMD hardware (GPUs, DSPs, ...).
% \end{itemize}
% \item LCA is inherently robust, is ``parameter free'' once setup with a well designed window set, and can often be implemented with minor modifications to already existing DAS beamformers.
% 
% The angular resolution and contrast in active sonar images depend on the beamformer's ability to receive signals from directions of interest, while suppressing noise and interference emanating from other directions. For sonar arrays, this is achieved by applying weights to the array channels.
% 
% While classical beamformers use predefined windows, adaptive beamformers estimate the optimal window by analytical evaluation of the data. The minimum variance (MV) beamformer, for instance, calculates the set of weights that minimises the variance of the beamformer's output. 
% 
We have studied the low complexity adaptive (LCA) beamformer in active sonar imaging. LCA can be viewed as either a simplification of the minimum variance distortionless response (MVDR) beamformer, or as an adaptive extension to the delay and sum (DAS) beamformer. While both LCA and MVDR attempt to minimize the power of noise and interference in the image, MVDR tries to find the optimal solution by estimating and inverting a spatial covariance matrix, while LCA iterates through a predefined set of sensor weights (or window) and selects the closest to optimal one.

@@

Since LCA operates with the same optimization criterion, we will be using approximately a factor 8 lateral oversampling for both beamformers. The images appear more visually pleasing and detailed up to 8. We consider an absolute minimum to be 2, due to the non-linear nature of delaying pixels that isn't strictly in the far field, and of displaying the absolute value of the pixels on a decibel scale. In \mediaII{} we visualize the effect of changing the lateral oversampling factor on a set of LCA images. The displayed images are all upinterpolated to the same size, but the detail level varies. 
% 
% \todo{Need $M$ in the image sector. Why???}

@@

\mediaI{} shows how \Fig{mvdr_selected_windows} changes as a function of the MVDR subarray length $L$ and temporal averaging $K$. It demonstrates that at $L=1$ the window responses are rectangular in all areas in the image. At $L=2$ we can observe that pixels in highlight are processed with windows with slightly varying amplitude responses, but with the phase being 0$^\circ$ in the illuminated seafloor sector. Yet MVDR is able to greatly suppress noise, which seems to emphasize the importance varying the zero locations slightly to suppress noise. The media file also shows MVDR can be run with subarray sizes $L\in[M/2,5M/8]$, but only with temporal averaging $K=1$ or above.

@@

The LCA images in \Fig{oversampling_mosaic_bounds} demonstrate the effect of adjusting the upper bound of $\beta$ and $\phi$. As suggested by \Fig{mvdr_selected_windows} we attempted first to use $\beta\in[0,2]$. While this resulted in much improved sidelobe suppression, there was also a slight improvement up to $\beta\in[0,5]$. This is likely needed to compensate for the increased sidelobe levels caused by steering. No noticeable difference were observed for higher values than 5. The images also demonstrate that increasing the steering $\phi$ improves the resolution of the strong scatterers. For this sonar the Rayleigh resolution is $\Delta=0.0125$ radians (0.72$^\circ$), thus a point scatterer imaged at 19\,m with a rectangular DAS would have a lateral FWHM of $0.125\cdot19\,\text{m}\cdot0.89=21\,\text{cm}$. This is close to the FWHM of the anchor in the DAS image in \Fig{oversampling_mosaic_bounds}, which means that the acoustic fingerprint of the anchor is similar to a point source. This suggests the anchor reflections to be specular and that we view its rounded side. As we increase steering to either 50\% or 100\% of the -3dB width, the FWHM drops to 73\% or 53\% to that of the rectangular window, respectively.

>>>>

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
% 
% \begin{itemize}
% \item Dynamic weights $\Rightarrow$ greater contrast/resolution potential
% \item But:
% \begin{itemize}
% \item Active system $\Rightarrow$ noise and signal correlated $\Rightarrow$ adaptive beamformers no longer robust.
% \item Subarray averaging decorrelate the noise, but in the process we sacrifice resolution.
% \item Computationally complex O($M^3$) due to inversion of a spatial covariance matrix. LCA is only O($MW$). Even if Capon subspace/beamspace versions exist, they remain more complex than the LCA, and is not as ideally suited for implementation of SIMD hardware (GPUs, DSPs, ...).
% \end{itemize}
% \item LCA is inherently robust, is ``parameter free'' once setup with a well designed window set, and can often be implemented with minor modifications to already existing DAS beamformers.
% 
% The angular resolution and contrast in active sonar images depend on the beamformer's ability to receive signals from directions of interest, while suppressing noise and interference emanating from other directions. For sonar arrays, this is achieved by applying weights to the array channels.
% 
% While classical beamformers use predefined windows, adaptive beamformers estimate the optimal window by analytical evaluation of the data. The minimum variance (MV) beamformer, for instance, calculates the set of weights that minimises the variance of the beamformer's output. 
% 
We have studied the low complexity adaptive (LCA) beamformer in active sonar imaging. LCA can be viewed as either a simplification of the minimum variance distortionless response (MVDR) beamformer, or as an adaptive extension to the delay and sum (DAS) beamformer. While both LCA and MVDR attempt to minimize the power of noise and interference in the image, MVDR tries find an optimal way of doing it by estimating and inverting the array's spatial covariance matrix, while LCA selects the best of a  predefined set of sensor window functions.

@@

Since LCA operates with the same optimization criterion, we will be using approximately a factor 8 lateral oversampling for both beamformers. The images appear more visually pleasing and detailed up to 8. We consider an absolute minimum to be 2, due to the non-linear nature of delaying pixels that isn't strictly in the far field, and of displaying the absolute value of the pixels on a decibel scale. In \mediaI{} we visualize the effect of changing the lateral oversampling factor on a set of LCA images. The displayed images are all upinterpolated to the same size, but the detail level varies. 
% 
% \todo{Need $M$ in the image sector. Why???}

@@

\mediaII{} shows how \Fig{mvdr_selected_windows} changes as a function of the MVDR subarray length $L$ and temporal averaging $K$. It demonstrates that at $L=1$ the window responses are rectangular in all areas in the image. At $L=2$ we can observe that pixels in highlight are processed with windows with slightly varying amplitude responses, but with the phase being 0$^\circ$ in the illuminated seafloor sector. Yet MVDR is able to greatly suppress noise, which seems to emphasize the importance varying the zero locations slightly to suppress noise. The media file also shows MVDR can be run with subarray sizes $L\in[M/2,5M/8]$, but only with temporal averaging $K=1$ or above.

@@

The LCA images in \Fig{oversampling_mosaic_bounds} demonstrate the effect of adjusting the upper bound of $\beta$ and $\phi$. As suggested by \Fig{mvdr_selected_windows} we attempted first to use $\beta\in[0,2]$. While this resulted in much improved sidelobe suppression, there was also a slight improvement up to $\beta\in[0,5]$. This is likely needed to compensate for the increased sidelobe levels caused by steering. No noticeable difference were observed for higher values than 5. The images also demonstrate that increasing the steering $\phi$ improves the resolution of the strong scatterers. For this sonar the Rayleigh resolution is $\Delta=0.0125$ radians (0.72$^\circ$), thus a point scatterer imaged at 19\,m with a rectangular DAS would have a lateral $\text{FWHM}\approx0.125\cdot19\,\text{m}\cdot0.89=21\,\text{cm}$. This is close to the FWHM of the anchor in the DAS image in \Fig{oversampling_mosaic_bounds}, which means that the acoustic fingerprint of the anchor is similar to a point source. This suggests the anchor reflections to be specular and that we view its rounded side. As we increase steering to either 50\% or 100\% of the -3dB width, the FWHM drops to 73\% or 53\% to that of the rectangular window, respectively.
